---
title: "Visualization"
subtitle: "part1"
output: github_document
---


```{r setup}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

scatterplots!
create my first scatterplot ever
```{r}
ggplot(weather_df,aes(x=tmin,y=tmax))+geom_point()
```

new approach, same plot
```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax))+geom_point()
```

save and edit a plot object
```{r}
weather_plot=
  weather_df %>% 
  ggplot(aes(x=tmin,y=tmax))
weather_plot+geom_point()
```


# advanced scatterplot

start with the same one and make it fancy

```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name))+geom_point()+geom_smooth(se=FALSE)
```

what about aes.. statement?

```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax))+geom_point(aes(color=name))+geom_smooth(se=FALSE)
```
facet
```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name,alpha=tmin))+geom_point()+geom_smooth(se=FALSE)+facet_grid(. ~name)
```
## some samll notes

how many geoms have to exist
```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name))+
  geom_smooth(se=FALSE)
```

you can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax))+
  geom_density2d()+
  geom_point(alpha=0.3)
```

## Univaribale plots
hitogram are really good
```{r}
weather_df %>% 
  ggplot(aes(x=tmin))+
  geom_histogram()
```

can we add color...
```{r}
weather_df %>% 
  ggplot(aes(x=tmin,fill=name))+
  geom_histogram(position="dodge")
```

lets try a new geometry
```{r}
weather_df %>% 
  ggplot(aes(x=tmin,fill=name))+
  geom_density(alpha=0.3)
```


what about box plot

```{r}
weather_df %>% 
  ggplot(aes(x=name,y=tmax))+
  geom_boxplot()
```


trendy plots

```{r}
weather_df %>% 
  ggplot(aes(x=name,y=tmin,fill=name))+
  geom_violin()+
  stat_summary(fun="median")
```


ridge plot

```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=name))+
  geom_density_ridges()
```

